---
title: "Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates"
author: Sununu Bah, Abbi Ross, Dylan Manfre, Ryan Mercado
date: 10/28/2022
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 



## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

elrich_contributions_2022 <- read_csv("data/elrich_contributions_2022.csv")
#This data set contains 2,981 rows and 13 columns. 
blair_contributions_2022 <- read_csv("data/blair_contributions_2022.csv")
#This data set contains 556 rows and 13 columns. 
elrich_contributions_2018 <- read_csv("data/elrich_contributions_2018.csv")
#This data set contains 4,010 rows and 13 columns. 
blair_contributions_2018 <- read_csv("data/blair_contributions_2018.csv")
#This data set contains 881 rows and 13 columns. 

dem_results_22 <- read_csv("data/dem_results_22.csv")
#This data set contains 32 rows and 11 columns. 
dem_results_18 <- read_csv("data/dem_results_18.csv")
#This data set contains 216 rows and 11 columns.
 

dem_precincts_22 <- read_csv("data/dem_precincts_22.csv")
#This data set contains 18,298 rows and 13 columns. 
dem_precincts_18 <- read_csv("data/dem_precincts_18.csv")
# This data set contains 27,791 rows and 11 columns.



# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")



# Clean required data and prepare for analysis if needed. 

elrich_contributions_2022 <- elrich_contributions_2022 %>% 
  clean_names() %>%
  select(-x13)

elrich_contributions_2018 <- elrich_contributions_2018 %>% 
  clean_names() %>% 
  select(-x13)

blair_contributions_2022 <- blair_contributions_2022 %>% 
  clean_names() %>% 
  select(-x13)

blair_contributions_2018 <- blair_contributions_2018 %>% 
  clean_names() %>%
  select(-x13)


dem_results_22 <- dem_results_22 %>% 
  clean_names() 

dem_results_18 <- dem_results_18 %>% 
  clean_names()


dem_precincts_22 <- dem_precincts_22 %>% 
  clean_names()

dem_precincts_18 <- dem_precincts_18 %>% 
  clean_names()

#Week Two Update: We  feel relatively comfortable with the data that we were able to find and download this week. Cleaning it was not that time consuming and it seems that we have all of the information we will need, other than the 2022 county results data and some iffy things surrounding mail/absentee ballots. Our first two questions are about self-funding and top contributors, which at this point do not seem like they should be an issue. Our third question is supposed to look at early voting and mail in votes. Right now it looks like the second half of the problem could be an issue. If so, we can tweak the question to focus solely on early votes and dive deeper into that. Our other questions are about campaign finances going toward ads and precinct level data, both of which we should be good to go on. The data sets we are analyzing are for two filing periods, 2022 and 2018. We chose these years because Elrich and Blair ran for the same position in both. The 2022 and 2018 campaign finance data sets are limited to any contributions made during that filing year. Other than the one data set not including all of the information we need, we're not seeing any major reasons that won't be able to complete our work. We did see where we will have to do something similar to our first project, where the state is identified by acronym instead of it's full name, so we will have to change that.Ryan did try to create a 2022 county result sheet, but we cannot get it to load.



```

## Sentences to Engineer
In this notebook, we are engineering five questions that we proposed after gathering blair and elrich contribution and election results data.

### Question 1

* **Question**: How much money did David Blair and Marc Elrich fund themselves in 2022 vs. 2018? Blair is a businessman and millionaire, so it makes sense that he’s funding his own campaign. But how much exactly is he putting into it, and how does it compare to the past election and to how much Elrich’s campaign is raising?

* **Analysis summary**:   

```{r}
blair_contributions_2022 %>%
  mutate(contributor_name = case_when(
  contributor_name == "BLAIR  DAVID  THOMAS" ~ "Blair  David  Thomas",
  TRUE ~ contributor_name
  )) %>% 
filter(contributor_name == "Blair  David  Thomas") %>%
group_by(contributor_name) %>%
summarize(total_blair = sum(contribution_amount))
```
```{r}
blair_contributions_2018 %>%
  mutate(contributor_name = case_when(
  contributor_name == "BLAIR  DAVID  THOMAS" ~ "Blair  David  Thomas",
  TRUE ~ contributor_name
  )) %>% 
filter(contributor_name == "Blair  David  Thomas") %>%
group_by(contributor_name) %>%
summarize(total_blair = sum(contribution_amount))
```
```{r}
elrich_contributions_2022 %>%
  filter(contributor_name == "Elrich  Marc") %>% 
  group_by(contributor_name) %>% 
  summarize(total_blair = sum(contribution_amount))
```
```{r}
elrich_contributions_2018 %>%
  filter(contributor_name == "Elrich  Marc") %>% 
  group_by(contributor_name) %>% 
  summarize(total_blair = sum(contribution_amount))
```
### Question 2

* **Question**: Who were the top 5 contributors to Blair and Elrich in 2022? How about in 2018? What are their connections to the candidates?

* **Analysis summary**:  

```{r}
top_blair_contribs_22 <- blair_contributions_2022 %>% 
  mutate(contributor_name = case_when(
    contributor_name == "BLAIR  DAVID  THOMAS" ~ "Blair  David  Thomas",
    TRUE ~ contributor_name
  )) %>% 
  group_by(contributor_name) %>% 
  summarize(total = sum(contribution_amount)) %>% 
  arrange(desc(total)) %>% 
  head(5)
top_blair_contribs_22
```

```{r}
top_blair_contribs_18 <- blair_contributions_2018 %>%
  mutate(contributor_name = case_when(
    contributor_name == "BLAIR  DAVID  THOMAS" ~ "Blair  David  Thomas",
    TRUE ~ contributor_name
  )) %>% 
  group_by(contributor_name) %>% 
  summarize(total = sum(contribution_amount)) %>% 
  arrange(desc(total)) %>% 
  head(5)
top_blair_contribs_18
```

```{r}
top_elrich_contribs_22 <- elrich_contributions_2022 %>% 
  group_by(contributor_name) %>% 
  summarize(total = sum(contribution_amount)) %>% 
  arrange(desc(total)) %>% 
  head(6)
top_elrich_contribs_22
```

```{r}
top_elrich_contribs_18 <- elrich_contributions_2018 %>% 
  group_by(contributor_name) %>% 
  summarize(total = sum(contribution_amount)) %>% 
  arrange(desc(total)) %>% 
  head(6)
top_elrich_contribs_18
```


### Question 3

* **Quesion**: David Blair got more early/election day votes in 2022. Elrich got more Mail-in votes in 2022. How did that compare to 2018? People want to know how Marc Elrich won both times, first by 72 voters in 2018 and then 32 in 2022. Where did Marc do well in the three categories: Early voting, election day, and mail-in votes. Did this sway the result at all?

* **Analysis summary**:   

```{r}
```

### Question 4

* **Question**: Which parts of the county voted for Elrich and which voted for Blair based on precinct-level voting? Are there differences in demographics of those areas? What about income?

* **Analysis summary**:   

```{r}
```


### Question 5

* **Question**: Where did David Blair and Marc Elrich spend their money on campaign finance resources? Ie: Meta advertisements, yard signs, TV, consulting? This would be interesting to know since Blair made a very big concerted effort to build his profile during the campaign while Elrich relied on incumbency and mainly TV ads to help him. 

* **Analysis summary**:   

```{r}
```
