---
title: "Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates"
author: Sununu Bah, Abbi Ross, Dylan Manfre, Ryan Mercado
date: 10/28/2022
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates](https://cnsmaryland.org/2022/03/03/millions-in-out-of-state-donations-help-fuel-high-profile-maryland-democratic-governor-candidates/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

elrich_contributions_2022 <- read_csv("data/elrich_contributions_2022.csv")
blair_contributions_2022 <- read_csv("data/blair_contributions_2022.csv")

elrich_contributions_2018 <- read_csv("data/elrich_contributions_2018.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")



# Clean required data and prepare for analysis if needed. 

elrich_contributions_2022 <- elrich_contributions_2022 %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date)) %>% 
  select(-x13)

elrich_contributions_2018 <- elrich_contributions_2018 %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date)) %>% 
  select(-x13)

blair_contributions_2022 <- blair_contributions_2022 %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date)) %>% 
  select(-x13)


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Question 1

* **Sentence text**: 

* **Analysis summary**:   

```{r}
```


### Question 2

* **Sentence text**: 

* **Analysis summary**:   

```{r}
```

### Question 3

* **Sentence text**: 

* **Analysis summary**:   

```{r}
```

### Question 4

* **Sentence text**: 

* **Analysis summary**:   

```{r}
```


### Question 5

* **Sentence text**: 

* **Analysis summary**:   

```{r}
```
