---
title: "Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates"
author: Sununu Bah, Abbi Ross, Dylan Manfre, Ryan Mercado
date: 10/28/2022
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates](https://cnsmaryland.org/2022/03/03/millions-in-out-of-state-donations-help-fuel-high-profile-maryland-democratic-governor-candidates/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

moore_contributions <- read_csv("data/moore_contributions.csv")

perez_contributions <- read.csv("data/perez_contributions.csv")

king_contributions <- read.csv("data/king_contributions.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")



# Clean required data and prepare for analysis if needed. 
moore_contributions <- moore_contributions %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date)) %>% 
  select(-x13)
glimpse(moore_contributions)

moore_contributions <- moore_contributions %>%
    mutate(contributor_state = case_when(
    str_detect(contributor_address, " AL ") ~ "Alabama",
    str_detect(contributor_address, " AK ") ~  "Alaska",
    str_detect(contributor_address, " AZ ") ~  "Arizona",
    str_detect(contributor_address, " AR ") ~  "Arkansas",
    str_detect(contributor_address, "5 Regent Dr Queanbeyan AS") ~  "American Samoa",
    str_detect(contributor_address, " CA ") ~  "California",
    str_detect(contributor_address, " CO ") ~ "Colorado",
    str_detect(contributor_address, " CT ") ~  "Connecticut",
    str_detect(contributor_address, " DE ") ~  "Delaware",
    str_detect(contributor_address, " FL ") ~  "Florida",
    str_detect(contributor_address, " GA ") ~  "Georgia",
    str_detect(contributor_address, " HI ") ~ "Hawaii",
    str_detect(contributor_address, " ID ") ~  "Idaho",
    str_detect(contributor_address, " IL ") ~  "Illinois",
    str_detect(contributor_address, " IN ") ~  "Indiana",
    str_detect(contributor_address, " IA ") ~  "Iowa",
    str_detect(contributor_address, " KS ") ~ "Kansas",
    str_detect(contributor_address, " KY ") ~  "Kentucky",
    str_detect(contributor_address, " LA ") ~  "Louisiana",
    str_detect(contributor_address, "13 Church Rd Flat 4 London UK") ~ "London",
    str_detect(contributor_address, " ME ") ~  "Maine",
    str_detect(contributor_address, " MD ") ~ "Maryland",
    str_detect(contributor_address, " MA ") ~  "Massachusetts",
    str_detect(contributor_address, " MI ") ~  "Michigan",
    str_detect(contributor_address, " MN ") ~ "Minnesota",
    str_detect(contributor_address, " MS ") ~  "Mississippi",
    str_detect(contributor_address, " MO ") ~  "Missouri",
    str_detect(contributor_address, " MT ") ~  "Montana",
    str_detect(contributor_address, " NE ") ~  "Nebraska",
    str_detect(contributor_address, " NV ") ~ "Nevada",
    str_detect(contributor_address, " NH ") ~ "New Hampshire",
    str_detect(contributor_address, " NJ ") ~ "New Jersey",
    str_detect(contributor_address, " NM ") ~ "New Mexico",
    str_detect(contributor_address, " NY ") ~  "New York",
    str_detect(contributor_address, " NC ") ~ "North Carolina",
    str_detect(contributor_address, " ND ") ~ "North Dakota",
    str_detect(contributor_address, " OH ") ~ "Ohio",
    str_detect(contributor_address, " OK ") ~ "Oklahoma", 
    str_detect(contributor_address, " OR ") ~ "Oregon", 
    str_detect(contributor_address, " PA ") ~ "Pennsylvania",
    str_detect(contributor_address, " RI ") ~ "Rhode Island",
    str_detect(contributor_address, "371 Beach Road Apt 11-04 Singapore SN") ~ "Singapore",
    str_detect(contributor_address, " SC ") ~ "South Carolina",
    str_detect(contributor_address, " SD ") ~ "South Dakota",
    str_detect(contributor_address, " TN ") ~ "Tennessee",
    str_detect(contributor_address, "95 Lower Way Thatcham UK") ~ "Thatcham",
    str_detect(contributor_address, " TX ") ~ "Texas",
    str_detect(contributor_address, " UT ") ~ "Utah",
    str_detect(contributor_address, " VT ") ~ "Vermont",
    str_detect(contributor_address, " VA ") ~ "Virginia",
    str_detect(contributor_address, " WA ") ~ "Washington",
    str_detect(contributor_address, " WV ") ~ "West Virginia",
    str_detect(contributor_address, " WI ") ~ "Wisconsin",
    str_detect(contributor_address, " WY ") ~ "Wyoming",
    str_detect(contributor_address, " DC ") ~ "Washington DC",
    TRUE ~ contributor_address
  ))



perez_contributions <- perez_contributions %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date)) %>% 
  select(-x)
glimpse(perez_contributions)

perez_contributions <- perez_contributions %>%
    mutate(contributor_state = case_when(
    str_detect(contributor_address, " AL ") ~ "Alabama",
    str_detect(contributor_address, " AK ") ~  "Alaska",
    str_detect(contributor_address, " AZ ") ~  "Arizona",
    str_detect(contributor_address, " AR ") ~  "Arkansas",
    str_detect(contributor_address, "5 Regent Dr Queanbeyan AS") ~  "American Samoa",
    str_detect(contributor_address, " CA ") ~  "California",
    str_detect(contributor_address, " CO ") ~ "Colorado",
    str_detect(contributor_address, " CT ") ~  "Connecticut",
    str_detect(contributor_address, " DE ") ~  "Delaware",
    str_detect(contributor_address, " FL ") ~  "Florida",
    str_detect(contributor_address, " GA ") ~  "Georgia",
    str_detect(contributor_address, " HI ") ~ "Hawaii",
    str_detect(contributor_address, " ID ") ~  "Idaho",
    str_detect(contributor_address, " IL ") ~  "Illinois",
    str_detect(contributor_address, " IN ") ~  "Indiana",
    str_detect(contributor_address, " IA ") ~  "Iowa",
    str_detect(contributor_address, " KS ") ~ "Kansas",
    str_detect(contributor_address, " KY ") ~  "Kentucky",
    str_detect(contributor_address, " LA ") ~  "Louisiana",
    str_detect(contributor_address, " ME ") ~  "Maine",
    str_detect(contributor_address, " MD ") ~ "Maryland",
    str_detect(contributor_address, " MA ") ~  "Massachusetts",
    str_detect(contributor_address, " MI ") ~  "Michigan",
    str_detect(contributor_address, " MN ") ~ "Minnesota",
    str_detect(contributor_address, " MS ") ~  "Mississippi",
    str_detect(contributor_address, " MO ") ~  "Missouri",
    str_detect(contributor_address, " MT ") ~  "Montana",
    str_detect(contributor_address, " NE ") ~  "Nebraska",
    str_detect(contributor_address, " NV ") ~ "Nevada",
    str_detect(contributor_address, " NH ") ~ "New Hampshire",
    str_detect(contributor_address, " NJ ") ~ "New Jersey",
    str_detect(contributor_address, " NM ") ~ "New Mexico",
    str_detect(contributor_address, " NY ") ~  "New York",
    str_detect(contributor_address, " NC ") ~ "North Carolina",
    str_detect(contributor_address, " ND ") ~ "North Dakota",
    str_detect(contributor_address, " OH ") ~ "Ohio",
    str_detect(contributor_address, " OK ") ~ "Oklahoma", 
    str_detect(contributor_address, " OR ") ~ "Oregon", 
    str_detect(contributor_address, " PA ") ~ "Pennsylvania",
    str_detect(contributor_address, " PR ") ~ "Puerto Rico",
    str_detect(contributor_address, "1453 Calle Salud") ~ "Puerto Rico",
    str_detect(contributor_address, " RI ") ~ "Rhode Island",
    str_detect(contributor_address, " SC ") ~ "South Carolina",
    str_detect(contributor_address, " SD ") ~ "South Dakota",
    str_detect(contributor_address, " TN ") ~ "Tennessee",
    str_detect(contributor_address, " TX ") ~ "Texas",
    str_detect(contributor_address, " UT ") ~ "Utah",
    str_detect(contributor_address, " VT ") ~ "Vermont",
    str_detect(contributor_address, " VA ") ~ "Virginia",
    str_detect(contributor_address, " WA ") ~ "Washington",
    str_detect(contributor_address, " WV ") ~ "West Virginia",
    str_detect(contributor_address, " WI ") ~ "Wisconsin",
    str_detect(contributor_address, " WY ") ~ "Wyoming",
    str_detect(contributor_address, " DC ") ~ "Washington DC",
    TRUE ~ contributor_address
  ))

king_contributions <- king_contributions %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date)) %>% 
  select(-x)
glimpse(king_contributions)


king_contributions <- king_contributions %>%
  mutate(contributor_state = case_when(
    str_detect(contributor_address, " AL ") ~ "Alabama",
    str_detect(contributor_address, " AK ") ~  "Alaska",
    str_detect(contributor_address, " AZ ") ~  "Arizona",
    str_detect(contributor_address, " AR ") ~  "Arkansas",
    str_detect(contributor_address, " CA ") ~  "California",
    str_detect(contributor_address, "208-6100 6th Avenue Whitehorse  YT CA") ~ "Canada",
    str_detect(contributor_address, " CO ") ~ "Colorado",
    str_detect(contributor_address, " CT ") ~  "Connecticut",
    str_detect(contributor_address, " DE ") ~  "Delaware",
    str_detect(contributor_address, " FL ") ~  "Florida",
    str_detect(contributor_address, " GA ") ~  "Georgia",
    str_detect(contributor_address, " HI ") ~ "Hawaii",
    str_detect(contributor_address, "27 C S BAY TOWERS 59 SOUTH BAY Rd Repulse Bay HK") ~ "Hong Kong",
    str_detect(contributor_address, " ID ") ~  "Idaho",
    str_detect(contributor_address, " IL ") ~  "Illinois",
    str_detect(contributor_address, " IN ") ~  "Indiana",
    str_detect(contributor_address, " IA ") ~  "Iowa",
    str_detect(contributor_address, " KS ") ~ "Kansas",
    str_detect(contributor_address, " KY ") ~  "Kentucky",
    str_detect(contributor_address, " LA ") ~  "Louisiana",
    str_detect(contributor_address, " ME ") ~  "Maine",
    str_detect(contributor_address, " MD ") ~ "Maryland",
    str_detect(contributor_address, " MA ") ~  "Massachusetts",
    str_detect(contributor_address, " MI ") ~  "Michigan",
    str_detect(contributor_address, " MN ") ~ "Minnesota",
    str_detect(contributor_address, " MS ") ~  "Mississippi",
    str_detect(contributor_address, " MO ") ~  "Missouri",
    str_detect(contributor_address, " MT ") ~  "Montana",
    str_detect(contributor_address, " NE ") ~  "Nebraska",
    str_detect(contributor_address, " NV ") ~ "Nevada",
    str_detect(contributor_address, " NH ") ~ "New Hampshire",
    str_detect(contributor_address, " NJ ") ~ "New Jersey",
    str_detect(contributor_address, " NM ") ~ "New Mexico",
    str_detect(contributor_address, " NY ") ~  "New York",
    str_detect(contributor_address, " NC ") ~ "North Carolina",
    str_detect(contributor_address, " ND ") ~ "North Dakota",
    str_detect(contributor_address, " OH ") ~ "Ohio",
    str_detect(contributor_address, " OK ") ~ "Oklahoma", 
    str_detect(contributor_address, " OR ") ~ "Oregon", 
    str_detect(contributor_address, " PA ") ~ "Pennsylvania",
    str_detect(contributor_address, " PR ") ~ "Puerto Rico",
    str_detect(contributor_address, " RI ") ~ "Rhode Island",
    str_detect(contributor_address, " SC ") ~ "South Carolina",
    str_detect(contributor_address, " SD ") ~ "South Dakota",
    str_detect(contributor_address, " TN ") ~ "Tennessee",
    str_detect(contributor_address, " TX ") ~ "Texas",
    str_detect(contributor_address, " UT ") ~ "Utah",
    str_detect(contributor_address, " VT ") ~ "Vermont",
    str_detect(contributor_address, " VA ") ~ "Virginia",
    str_detect(contributor_address, " WA ") ~ "Washington",
    str_detect(contributor_address, " WV ") ~ "West Virginia",
    str_detect(contributor_address, " WI ") ~ "Wisconsin",
    str_detect(contributor_address, " WY ") ~ "Wyoming",
    str_detect(contributor_address, " DC ") ~ "Washington DC",
    TRUE ~ contributor_address
  ))


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “We set the pace in fundraising — including raising more money in Maryland and having 73% of donations $100 or less — because this campaign has inspired thousands of people from all walks of life in Maryland and across the country who see the opportunity to make Wes Moore governor of Maryland and change the trajectories of families who have been left behind,” the campaign said in a statement."

* **Analysis summary**: Our data says that only about 62% of Moore's donations are $100 or less. We went up to the date that the article was published. I'm not sure why the data is off. We even went up to 12/31/2021 but the percent just went up 1%. We cleaned and sorted everything properly.  

```{r}
moore_contributions %>%
  filter(contribution_date <= "2022-03-02") %>%
  summarize(total_contributions = n(), total_contribution_amount = sum(contribution_amount))
```

```{r}
# Put code to reverse engineer sentence here
less_than_hundred_moore_contribs <- moore_contributions %>% 
  filter(contribution_amount < 100)
```

```{r}  
less_than_hundred_moore_contribs %>%
  filter(contribution_date <= "2022-03-02") %>%
  summarize(less_than_hundred_contribs = n(), less_than_hundred_amount = sum(contribution_amount))

  
# Display results of code below this codeblock

```

### Sentence 2 & 3

* **Sentence text**: “Moore, a U.S. Army veteran and bestselling author, built a campaign war chest that included over $2.3 million in out-of-state contributions. 

Tom Perez, a former leader of the Democratic National Committee, raised nearly $1.7 million in donations from outside Maryland.” 



* **Analysis summary**: Our data said that Moore had more than $2.3 million out-of-state contributions. 5,006 contributions for $2,403,216	to be exact.

Perez on the other hand, had more than $1.7 million out-of-state contributions. 2,180 contributions for $1,696,293 to be exact. 

January 12 was the only date that could get us as close to the numbers as possible. But the article said Moore had more than $2.3 million instead of more than $2.4 million. That's why I believe this date may not be the correct cut-off point or there's something else wrong that we're not sure of.


```{r}
moore_contributions %>%
  filter(contribution_date <= "2022-01-12") %>% 
  group_by(contributor_state) %>%
  summarise(total_state_amount = sum(contribution_amount),
            mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(total_state_amount))
```
```{r}
perez_contributions %>%
  filter(contribution_date <= "2022-01-12") %>% 
  group_by(contributor_state) %>%
  summarise(total_state_amount = sum(contribution_amount),
            mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(total_state_amount))
```

```{r}
out_of_state_moore_contributions <- moore_contributions %>% 
  filter(contribution_date <= "2022-01-12") %>% 
  filter(!str_detect(contributor_state, "Maryland"))
```

```{r}
out_of_state_moore_contributions %>%
  summarise(out_of_state_amount = sum(contribution_amount),
          mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(out_of_state_amount))

```

```{r}
out_of_state_perez_contributions <- perez_contributions %>%
  filter(contribution_date <= "2022-01-12") %>% 
  filter(!str_detect(contributor_state, "Maryland"))
```

```{r}
out_of_state_perez_contributions %>%
  summarise(out_of_state_amount = sum(contribution_amount),
          mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(out_of_state_amount))

```

### Sentence 4

* **Sentence text**: John B. King, former secretary of Education under President Barack Obama, raised $2.2 million in campaign funding last year.

* **Analysis summary**: Our code said the John B. King raised less than $2.2 million in campaign funding last year. Our code said that he raised $2,018,222 for the entire 2021 year to be exact.  

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

king_contributions <- king_contributions %>% 
filter(contribution_date <= "2021-12-31")
```

```{r}
king_contributions %>% 
summarize(total_2021 = sum(contribution_amount)) %>% 
arrange(desc(total_2021))

```

### Sentence 5

* **Sentence text**: Moore and King collected over $100,000 each from Illinois.
* **Analysis summary**: Our code matches this. Both Moore and King collected over $100,000 each from Illinois. Moore collected $160,075.6 while King collected $118,520.2. 


```{r}
moore_contributions %>% 
  filter(contributor_state == "Illinois") %>% 
  group_by(contributor_state) %>% 
  summarise(total_amount = sum(contribution_amount))
```

```{r}
king_contributions %>% 
  filter(contributor_state == "Illinois") %>% 
  group_by(contributor_state) %>% 
  summarise(total_amount = sum(contribution_amount))
```
-30-