---
title: "Millions in out-of-state donations help fuel high-profile Maryland Democratic governor candidates"
author: "Stephen Neukam"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

moore_contributions <- read_csv("data/moore_contributions.csv")
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
glimpse(moore_contributions)

moore_contributions <- moore_contributions %>% 
  clean_names() %>% 
  mutate(contribution_date = mdy(contribution_date))
glimpse(moore_contributions)
# Clean required data and prepare for analysis if needed. 




```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: Finding all of Moore's contributions that are less than $100 and seeing the top donors of $100 and less.
* **Analysis summary**: Wes Moore had 8,647 unique individuals donate less than $100. One of the individuals donated less than $100 82 times. Her name is Delores Parron-Ragland. She was a distinguished social worker who's career spans in the federal government, academia, and clinical practice. She's an outlier however. The top doners 28 to 44 contributions of less than $100 however. 
```{r}
# Put code to reverse engineer sentence here
less_than_hundred_moore_contribs <- moore_contributions %>% 
  filter(contribution_amount < 100)
```

```{r}  
less_than_hundred_moore_contribs <- less_than_hundred_moore_contribs %>%
  group_by(contributor_name) %>% 
  summarise(total = n(), amount = sum(contribution_amount)) %>% 
  arrange(desc(total))
less_than_hundred_moore_contribs
  
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-