---
title: "lab_03"
author: "Derek Willis"
date: "8/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
```

## Load data
```{r}
# Write code to load MD slice of FEC House candidate contribution data and call it md_contribs.
# The data is called lab_03.rds, and is in the data folder alongside lab_03.Rmd (this data notebook)
md_contribs <- read_rds("data/lab_03.rds")

```

## Answer questions

Q1: Which seven cities in Maryland had the largest _number_ of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1: 1. Baltimore: 6726 2. Bethesda: 5860 3. Silver Spring: 4659 4. Chevy Chase: 3260 5. Potomac: 2302 6. Rockville: 2213 7. Annapolis: 2001



```{r}
md_contribs %>% 
  group_by(contributor_city) %>% 
  summarize(total = n()) %>% 
  arrange(desc(total))
```

```{r}
# Write your code in this codeblock.
cleaned_city <- md_contribs %>%
  mutate(contributor_city = str_to_title(contributor_city)) %>% 
  mutate(contributor_city = case_when(
    str_detect(contributor_city,"^11509") ~ "Potomac",
    TRUE ~ contributor_city
  )) %>%
  mutate(contributor_city = case_when(
    str_detect(contributor_city, "^Pot") ~ "Potomac",
    TRUE ~ contributor_city
  )) %>% 
  mutate(contributor_city = case_when(
    str_detect(contributor_city,"^15600") ~ "Upper Marlboro",
    TRUE ~ contributor_city
  )) %>% 
  mutate(contributor_city = case_when(
    str_detect(contributor_city,"^15616") ~ "Bowie",
    TRUE ~ contributor_city
  )) %>% 
  mutate(contributor_city = case_when(
    str_detect(contributor_city, "^Beth") ~ "Bethesda",
    TRUE ~ contributor_city
  )) %>% 
  mutate(contributor_city = case_when(
    str_detect(contributor_city, "^Silver") ~ "Silver Spring", 
    TRUE ~ contributor_city
  )) %>% 
mutate(contributor_city = case_when(
  contributor_city == "Sliver Spring" ~ "Silver Spring",
  TRUE ~ contributor_city
)) %>% 
mutate(contributor_city = case_when(
  str_detect(contributor_city, "^Balit") ~ "Baltimore",
  TRUE ~ contributor_city
)) %>% 
  mutate(contributor_city = case_when(
    str_detect(contributor_city, "^Anap") ~ "Annapolis",
    TRUE ~ contributor_city
  )) %>% 
  mutate(contributor_city = case_when(
  str_detect(contributor_city, "^Anretired") ~ "Annapolis",
  TRUE ~ contributor_city
  )) %>% 
  mutate(contributor_city = case_when(
  contributor_city == "Annapolid" ~ "Annapolis",
  TRUE ~ contributor_city
  )) %>% 
    mutate(contributor_city = case_when(
    str_detect(contributor_city, "^Annn") ~ "Annapolis",
    TRUE ~ contributor_city
  ))


```

```{r}
cleaned_city %>% 
  group_by(contributor_city) %>% 
  summarize(total = n()) %>% 
  arrange(desc(total))
```



Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). Where is that zip code and what kind of area is it? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: 20854 = Potomac. Potomac is one of the wealthiest and most educated cities in the entire country. JAMIE RASKIN FOR CONGRESS is the largest recipient. 3 of the top 5 recipients are Maryland candidates: Jamie Raskin, Aruna Miller, Heather Mizeur.



```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
cleaned_city %>% 
  group_by(contributor_zip, contributor_city) %>% 
  summarize(total_amount = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total_amount))
```
```{r}
# Second, use the top ZIP code in the dataframe you just created and create a new dataframe that shows the committee that got the most money from that ZIP code.
cleaned_potomac <- cleaned_city %>% 
  group_by(contributor_zip, committee_name ) %>% 
  summarize(total_amount = sum(contribution_receipt_amount)) %>% 
  filter(str_detect(contributor_zip, "20854")) %>% 
  arrange(desc(total_amount)) 

  cleaned_potomac

```
```{r}
cleaned_potomac %>% 
  group_by(committee_name) %>% 
  summarize(total = sum(total_amount)) %>% 
  arrange(desc(total))
```



Q3: Let's look at top individual donors to House campaigns. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. One of the top few donors has a very strong connection to the U.S. House - who is that and what does he do? How does his donation pattern differ from the second-largest donor, and what can you say about that second-largest donor?

Feel free to add any interesting "color" (little interesting details) you find in your research.  

A3: The top donor is David Trone who is a democrat member of the United States House of Representatives from Maryland's 6th congressional district. He is the founder and co-owner of Total Wine & More, the largest privately owned beer, wine and spirits retailer in the United States. He set the record for the most expensive self-funded House campaigns, which he lost to Chris Van Hollen (Maryland, 8th district). 

He donated $86,894.43 more than second-largest donor, Juanita Mason and did so with 3,017 less contributions. The reason for that is because the lowest he was donating was $1,000 and the average he was donating was $2,689 and the most he was donating was $5800. Juanita on the other hand had a negative minimum amount of -$405, which would be interesting to explore. The average that she was donating was around $28 and the most she donated was $405. 



```{r}
# Hint
# First, create a dataframe of top individual contributors
md_contribs %>%
    mutate(contributor_first_name = case_when(
    str_detect(contributor_first_name,"^REP.") ~ "DAVID",
    TRUE ~ contributor_first_name
    )) %>% 
  group_by(contributor_first_name, contributor_last_name) %>% 
  summarize(total_amount = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total_amount))


```

```{r}
# Second, use information from that dataframe to examine individual donations from the original md_contribs dataframe.
david_trone_contribs <- md_contribs %>% 
  filter(str_detect(contributor_first_name, "DAVID")) %>% 
  filter(str_detect(contributor_last_name, "TRONE")) %>%
  
    mutate(contributor_first_name = case_when(
    str_detect(contributor_first_name,"^REP.") ~ "DAVID",
    TRUE ~ contributor_first_name
  )) %>% 
  group_by(contributor_first_name, contributor_last_name) %>% 
  summarize(total_contribs = n(), 
            total_amount = sum(contribution_receipt_amount), 
            mean_amount = mean(contribution_receipt_amount),
            median_amount = median(contribution_receipt_amount),
            min_amount = min(contribution_receipt_amount),
            max_amount = max(contribution_receipt_amount)) 
  david_trone_contribs

```
```{r}
juanita_mason_contribs <- md_contribs %>% 
  filter(str_detect(contributor_first_name, "JUANITA")) %>% 
  filter(str_detect(contributor_last_name, "MASON")) %>%
  
  group_by(contributor_first_name, contributor_last_name) %>% 
  summarize(total_contribs = n(), 
            total_amount = sum(contribution_receipt_amount), 
            mean_amount = mean(contribution_receipt_amount),
            median_amount = median(contribution_receipt_amount),
            min_amount = min(contribution_receipt_amount),
            max_amount = max(contribution_receipt_amount)) 
  juanita_mason_contribs
  david_trone_contribs

```
