---
title: "Chris Van Hollen"
author: "Lisa Woelfl"
date: "10/11/22"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are looking at FEC data concerning Chris Van Hollen's senator race campaign.

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
# Clean required data and prepare for analysis if needed. 
receipts_hollen <- read_csv("data/receipts_schedule_2022.csv")
disburs_hollen <- read_csv("data/disbursements_schedule_2022.csv")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Question 1

* **Question**: Where is Van Hollen's money coming from? Who is donating the most? Who are those people/organizations?

* **Analysis summary**: 

```{r}
head(receipts_hollen)
```

```{r}
total_donations_hollen <- receipts_hollen %>% 
  group_by(contributor_name) %>% 
  summarise(total = sum(contribution_receipt_amount)) %>% 
  arrange(desc(total))

total_donations_hollen
```

```{r}
totalsum_donations_hollen <- receipts_hollen %>%
  group_by(report_year) %>% 
  summarise(total = sum(contribution_receipt_amount))

totalsum_donations_hollen
```


```{r}
count_donations_hollen <- receipts_hollen %>% 
  group_by(contributor_name) %>% 
  summarise(count_donations = n()) %>% 
  arrange(desc(count_donations))

count_donations_hollen
```

```{r}
receipts_hollen %>% 
  filter(contributor_name == "PARRON-RAGLAND, DELORES L")
```








### Question 2

* **Question**: Where is Van Hollen spending his money? How much? Where is most of his money going? Who are those people/organizations?

```{r}
totalsum_disburs_hollen <- disburs_hollen %>%
  group_by(report_year) %>% 
  summarise(total = sum(disbursement_amount))

totalsum_disburs_hollen
```

```{r}
count_disburs_hollen <- disburs_hollen %>% 
  group_by(recipient_name) %>% 
  summarise(count_disburs = n()) %>% 
  arrange(desc(count_disburs))

count_disburs_hollen
```

```{r}
indi_disburs_hollen <- disburs_hollen %>% 
  group_by(recipient_name) %>% 
  filter(entity_type_desc == "INDIVIDUAL") %>% 
  summarise(total = sum(disbursement_amount)) %>% 
  arrange(desc(total))

indi_disburs_hollen
```
Screen Strategies is a TV advertising company that specializes in developing TV advertising for Democratic candidates 

```{r}
total_disburs_hollen <- disburs_hollen %>% 
  group_by(recipient_name) %>% 
  summarise(total = sum(disbursement_amount)) %>% 
  arrange(desc(total))

total_disburs_hollen
```










