---
title: "Chris Van Hollen"
author: "Sununu Bah"
date: "10/11/22"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are looking at FEC data concerning Chris Van Hollen's senator race campaign.

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
# Clean required data and prepare for analysis if needed. 
receipts_hollen <- read_csv("data/receipts_schedule_2022.csv")
disburs_hollen <- read_csv("data/disbursements_schedule_2022.csv")
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Question 1

***Question 1*
What is the total number/amount of out-of-state and in-state contributions?
```{r}
receipts_hollen %>% 
  group_by(contributor_state) %>%
  summarise(total_amount = sum(contribution_receipt_amount),
            mean_amount= mean(contribution_receipt_amount),
          number_of_contribs=n())%>%
arrange(desc(total_amount))
```
```{r}
receipts_hollen %>% 
  summarise(total_amount = sum(contribution_receipt_amount),
            mean_amount= mean(contribution_receipt_amount),
          number_of_contribs=n())%>%
arrange(desc(total_amount))
```

```{r}
out_of_state_receipts <- receipts_hollen %>% 
  filter(!str_detect(contributor_state, "MD"))
```

```{r}
out_of_state_receipts %>% 
  summarise(out_of_state_amount = sum(contribution_receipt_amount),
            mean_amount= mean(contribution_receipt_amount),
          number_of_contribs=n())%>%
arrange(desc(out_of_state_amount))
```



***Question 2*** What is the total number/amount of out-of-state and in-state contributions?
```{r}
disburs_hollen %>% 
  group_by(recipient_state) %>%
  summarise(total_amount = sum(disbursement_amount),
            mean_amount= mean(disbursement_amount),
          number_of_contribs=n())%>%
arrange(desc(total_amount))

```
```{r}
disburs_hollen %>% 
summarise(total_amount = sum(disbursement_amount),
            mean_amount= mean(disbursement_amount),
          number_of_contribs=n())%>%
arrange(desc(total_amount))
```

```{r}
out_of_state_disbursements <- disburs_hollen %>% 
  filter(!str_detect(recipient_state, "MD"))
```

```{r}
out_of_state_disbursements %>% 
summarise(out_of_state_amount = sum(disbursement_amount),
            mean_amount= mean(disbursement_amount),
          number_of_contribs=n())%>%
arrange(desc(out_of_state_amount))
```








